---
title: "Directed Participation Networks - Fig. 3"
author: "Mary Fisher"
date: "10/9/2020"
output: html_document
---


# Description

Graph the directed networks which show flow of Dungeness crab vessels between fisheries in the early season of the 2015 --> early season of 2016. 

For more details on how directed vessel networks are built and interpreted, see the [Guide to Directed Networks](https://github.com/mfisher5/ParticipationNetworks/blob/master/doc/Guide_To_Directed_Networks.md) on Github.


```{r setup, include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("igraph")) {install.packages("igraph")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("ggraph")) {install.packages("ggraph")}
if(!require("grid")) {install.packages("grid")}
if(!require("gridExtra")) {install.packages("gridExtra")}
```
<br>

And calls the following functions:
```{r}
source("R/rename_network_nodes.R")
source("R/graph_directed_network_early.R")
source("R/color_network_nodes.R")
```
<br>

# User Inputs 

Select your directories.
```{r get_dir}
## location of igraph objects
indir <- 'data/networks/vessel_flow'
```
<br>

Identify the port groups that you would like to produce networks for. The port groups vector should include only those port groups which have associated `igraph` objects. Also list out the full names of the ports, so that names are used instead of codes on the figure.
```{r}
myports <- c("CCA", "ERA", "BGA", "BDA", "SFA", "MRA", "MNA")
portnames <- c("Crescent City", "Eureka", "Fort Bragg", "Bodega Bay", "San Francisco", "Monterey Bay", "Morro Bay")
```
<br>



# Plot Directed Networks

A quick for-loop to save all of the graph objects into a list.
```{r}
graphs_list <- list()
list_names <- c()
i=1
for(p in myports){
  for(s in c("large","small")){
    tmpgraph <- readRDS(here::here(indir,paste0(p,"_closure_",s,"_v7_newVL.rds")))
    V(tmpgraph)$port = portnames[which(myports == p)]
    V(tmpgraph)$vsize = s
    tmpgraph <- rename_vertices(tmpgraph)
    graphs_list[[i]] <- tmpgraph
    list_names[i] <- paste0(p,"_",s)
    i = i + 1
  }
}
names(graphs_list) <- list_names
```
<br>

Apply the plotting function to all of the graphs in the list.
```{r}
myplots <- lapply(graphs_list, plot_directed_early)
```
<br>

Create Figure 3, with column and x axis titles.
```{r}
col.titles=c("Large","Small")
ylabel="2015                                      2016"
png(here::here("results/figures","Figure3.png"), width=1800,height=3000,res=200)
grid.arrange(grobs=lapply(c(1,2), function(i) {
  arrangeGrob(grobs=myplots[seq(i,14,by=2)], top=textGrob(col.titles[i],gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
}), ncol=2)
dev.off()
```
<br>

---

`r Sys.time() - script_start_time`







