---
title: "Directed Participation Networks - Fig. 3"
author: "Mary Fisher"
date: "10/9/2020"
output: html_document
---


# Description

Graph the directed networks which show flow of Dungeness crab vessels between fisheries in the early season of the 2015 --> early season of 2016. 

For more details on how directed vessel networks are built and interpreted, see the [Guide to Directed Networks](https://github.com/mfisher5/ParticipationNetworks/blob/master/doc/Guide_To_Directed_Networks.md) on Github.


```{r setup, include=FALSE}
if(!require("here")) {install.packages("here")}
library(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())

## start time for full script
script_start_time <- Sys.time()
```
<br>

This script requires the following packages. 
```{r packages, message=FALSE, warning=FALSE}
if(!require("tidyverse")) {install.packages("tidyverse")}
if(!require("igraph")) {install.packages("igraph")}
if(!require("ggplot2")) {install.packages("ggplot2")}
if(!require("ggraph")) {install.packages("ggraph")}
if(!require("grid")) {install.packages("grid")}
if(!require("ggimage")) {install.packages("ggimage")}
if(!require("gridExtra")) {install.packages("gridExtra")}
```
<br>

And calls the following functions:
```{r}
source("R/rename_network_nodes.R")
source("R/graph_directed_network_early.R")
source("R/color_network_nodes.R")
```
<br>

# User Inputs 

Select your directories.
```{r get_dir}
## location of igraph objects
indir <- 'data/networks/vessel_flow'
```
<br>

Identify the port groups that you would like to produce networks for. The port groups vector should include only those port groups which have associated `igraph` objects. Also list out the full names of the ports, so that names are used instead of codes on the figure.
```{r}
myports <- c("CCA", "ERA", "BGA", "BDA", "SFA", "MRA", "MNA")
portnames <- c("Crescent City", "Eureka", "Fort Bragg", "Bodega Bay", "San Francisco", "Monterey", "Morro Bay")
```
<br>



# Plot Directed Networks

A quick for-loop to save all of the graph objects into a list.
```{r}
graphs_list <- list()
list_names <- c()
i=1
for(p in myports){
  for(s in c("large","small")){
    tmpgraph <- readRDS(here::here(indir,paste0(p,"_closure_",s,"_v7_newVL.rds")))
    V(tmpgraph)$port = portnames[which(myports == p)]
    V(tmpgraph)$vsize = s
    tmpgraph <- rename_vertices(tmpgraph)
    graphs_list[[i]] <- tmpgraph
    list_names[i] <- paste0(p,"_",s)
    i = i + 1
  }
}
names(graphs_list) <- list_names
```
<br>

## With D. crab node labeled

Apply the plotting function to all of the graphs in the list.
```{r}
myplots <- lapply(graphs_list, plot_directed_early)
```
<br>

Create Figure 3, with column and x axis titles.
```{r}
col.titles=c("Large","Small")
ylabel="2015                                      2016"
png(here::here("results/figures","Figure3.png"), width=1800,height=3000,res=200)
grid.arrange(grobs=lapply(c(1,2), function(i) {
  arrangeGrob(grobs=myplots[seq(i,14,by=2)], top=textGrob(col.titles[i],gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
}), ncol=2)
dev.off()
```
<br>

# With D. crab node as icon
```{r}
plot_directed_early = function (g) {
  ####### for all graphs ######
  # save the port group
  p=unique(V(g)$port) # port group
  # color the network nodes
  V(g)$color <- vertex_color(g) # node colors
  # remove any nodes with size of 0 in 2015 (except for Dungeness crab)
  g <- delete.vertices(g, V(g)[which(V(g)$size==0 & V(g)$name != "DCRB_POT")])
  
  ## get node sizes, and reset 2015 d.crab node size to the 2014 participation
  dcrb_node <- which(V(g)$name=="DCRB_POT")
  vsizes=V(g)$size
  vsizes[dcrb_node] = V(g)$size14[dcrb_node]
  
  # to plot vertices with only self loops on the left side of the graph
  self_only <- c()
  self_g = delete.edges(g,E(g)[!(which_loop(g))])
  self_loops <- as_edgelist(self_g)[,1]
  
  # set the 'y' axis locations of each node. This also requires you to set the start and end of the edges. 
  node.y <- unlist(lapply(V(g)$name,FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1,ifelse(i %in% self_loops, 0.1,0))}))
  edge.y <- unlist(lapply(as_edgelist(g)[,1],FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1,ifelse(i %in% self_loops, 0.1,0))}))
  edge.yend <- unlist(lapply(as_edgelist(g)[,2],FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1,ifelse(i %in% self_loops, 0.1,0))}))
  node.y.lab <- unlist(lapply(V(g)$name,FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1.2,ifelse(i %in% self_loops, -0.5,-0.5))}))
  
  # crab node location
  crab_node <- data.frame(x=-0.5, y=1, crab_image=here::here('data/input', 'crab_icon_NWFSCgraphics.png'))
  # empty crab node common name so it isn't labeled on the graph
  V(g)$common_name[which(V(g)$common_name=="D.crab")] <- ""
  
  ####### small / large vessel graphs ######
  # save the size category
  s=unique(V(g)$vsize)
  s=ifelse(s=="large","Large","Small")
  # the graph will be labeled with the port group, for large vessels (on left of combined graph)
  if(s=="Large"){ylabel=p} else{ylabel=""}
  
  ### if none of the vertices have self-loops ###
  if(ecount(self_g) < 1){
    ggraph(g, 'igraph', algorithm = 'tree') + 
      geom_edge_diagonal(aes(y=edge.y,yend=edge.yend,width=sqrt((E(g)$weight))), 
                         color="grey85",end_cap = circle(0.5),arrow=arrow(length=unit(0.3,'cm'))) +
      geom_node_point(aes(y=node.y),color=factor(V(g)$color),
                      size=12*sqrt(vsizes)/sqrt(max(vsizes))) +
      geom_node_text(aes(y=node.y.lab,label = common_name), size=5) +
      geom_image(data=crab_node, aes(x=x,y=y,image=crab_image), size = 0.25, by="width") +
      scale_size_identity() +
      #annotate("text", x = 1.35, y = 0.5, label = paste0(s," Vessels"), size=4) +
      xlab(ylabel) +
      xlim(c(-2.7,2.7)) +
      # ylim(c(-0.05,1.1)) +
      theme_void() +
      theme(legend.position="none", 
            axis.title.y=element_text(angle=90,size=20,hjust=0.5),
            panel.border = element_rect(colour = "black", fill=NA, size=1)) +
      coord_flip() +
      scale_y_reverse(expand=expansion(add=c(0.5,1)))
    ### if 1+ vertices have self-loops ###
  } else{
    ggraph(g, 'igraph', algorithm = 'tree') + 
      geom_edge_diagonal(aes(y=edge.y,yend=edge.yend,width=sqrt((E(g)$weight))), 
                         color="grey85",end_cap = circle(0.5),arrow=arrow(length=unit(0.3,'cm'))) +
      geom_edge_loop(aes(y=edge.y,width=sqrt((E(g)$weight)), span=50,direction=0,strength=0.90), color="grey85") +
      geom_node_point(aes(y=node.y),color=factor(V(g)$color),
                      size=12*sqrt(vsizes)/sqrt(max(vsizes))) +
      geom_node_text(aes(y=node.y.lab,label = common_name), size=5) +
      geom_image(data=crab_node, aes(x=x,y=y,image=crab_image), size = 0.25, by="width") +
      scale_size_identity() +
      xlab(ylabel) +
      xlim(c(-2.7,2.7)) +
      # ylim(c(-0.09,1.15)) +
      theme_void() +
      theme(legend.position="none", 
            axis.title.y=element_text(angle=90,size=20,hjust=0.5),
            panel.border = element_rect(colour = "black", fill=NA, size=1)) +
      coord_flip() +
      scale_y_reverse(expand=expansion(add=c(0.15,1)))
  }
}
```
<br>


Apply the plotting function to all of the graphs in the list.
```{r}
myplots <- lapply(graphs_list, plot_directed_early)
```
<br>

Create Figure 3, with column and x axis titles.
```{r}
col.titles=c("Large","Small")
ylabel="2015                                      2016"
png(here::here("results/figures","Figure3_icon.png"), width=1800,height=3000,res=200)
grid.arrange(grobs=lapply(c(1,2), function(i) {
  arrangeGrob(grobs=myplots[seq(i,14,by=2)], top=textGrob(col.titles[i],gp=gpar(fontsize=22)), 
              bottom=textGrob(ylabel,gp=gpar(fontsize=18)),
              ncol=1)
}), ncol=2)
dev.off()
```


---

`r Sys.time() - script_start_time`



```{r}
crab_image <- readPNG(here::here('data/input', 'crab_icon_NWFSCgraphics.png'))
g <- graphs_list[[1]]
  
  ####### for all graphs ######
  # save the port group
  p=unique(V(g)$port) # port group
  # color the network nodes
  V(g)$color <- vertex_color(g) # node colors
  # remove any nodes with size of 0 in 2015 (except for Dungeness crab)
  g <- delete.vertices(g, V(g)[which(V(g)$size==0 & V(g)$name != "DCRB_POT")])
  
  ## get node sizes, and reset 2015 d.crab node size to the 2014 participation
  dcrb_node <- which(V(g)$name=="DCRB_POT")
  vsizes=V(g)$size
  vsizes[dcrb_node] = V(g)$size14[dcrb_node]
  
  # to plot vertices with only self loops on the left side of the graph
  self_only <- c()
  self_g = delete.edges(g,E(g)[!(which_loop(g))])
  self_loops <- as_edgelist(self_g)[,1]
  
  # set the 'y' axis locations of each node. This also requires you to set the start and end of the edges. 
  node.y <- unlist(lapply(V(g)$name,FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1,ifelse(i %in% self_loops, 0.1,0))}))
  edge.y <- unlist(lapply(as_edgelist(g)[,1],FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1,ifelse(i %in% self_loops, 0.1,0))}))
  edge.yend <- unlist(lapply(as_edgelist(g)[,2],FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1,ifelse(i %in% self_loops, 0.1,0))}))
  node.y.lab <- unlist(lapply(V(g)$name,FUN=function(i){ifelse(i %in% c("DCRB_POT",self_only), 1.2,ifelse(i %in% self_loops, -0.5,-0.5))}))
  
  # crab node location
  crab_node <- data.frame(x=-0.5, y=1, crab_image=here::here('data/input', 'crab_icon_NWFSCgraphics.png'))
  # empty crab node common name so it isn't labeled on the graph
  V(g)$common_name[which(V(g)$common_name=="D.crab")] <- ""
  
  ####### small / large vessel graphs ######
  # save the size category
  s=unique(V(g)$vsize)
  s=ifelse(s=="large","Large","Small")
  # the graph will be labeled with the port group, for large vessels (on left of combined graph)
  if(s=="Large"){ylabel=p} else{ylabel=""}
  
  ### if none of the vertices have self-loops ###
  if(ecount(self_g) < 1){
    ggraph(g, 'igraph', algorithm = 'tree') + 
      geom_edge_diagonal(aes(y=edge.y,yend=edge.yend,width=sqrt((E(g)$weight))), 
                         color="grey85",end_cap = circle(0.5),arrow=arrow(length=unit(0.3,'cm'))) +
      geom_node_point(aes(y=node.y),color=factor(V(g)$color),
                      size=12*sqrt(vsizes)/sqrt(max(vsizes))) +
      geom_node_text(aes(y=node.y.lab,label = common_name), size=5) +
      geom_image(data=crab_node, aes(x=x,y=y,image=crab_image), size=0.5) +
      #annotate("text", x = 1.35, y = 0.5, label = paste0(s," Vessels"), size=4) +
      xlab(ylabel) +
      xlim(c(-2.7,2.7)) +
      # ylim(c(-0.05,1.1)) +
      theme_void() +
      theme(legend.position="none", 
            axis.title.y=element_text(angle=90,size=20,hjust=0.5),
            panel.border = element_rect(colour = "black", fill=NA, size=1)) +
      coord_flip() +
      scale_y_reverse(expand=expansion(add=c(0.5,1)))
    ### if 1+ vertices have self-loops ###
  } else{
    ggraph(g, 'igraph', algorithm = 'tree') + 
      geom_edge_diagonal(aes(y=edge.y,yend=edge.yend,width=sqrt((E(g)$weight))), 
                         color="grey85",end_cap = circle(0.5),arrow=arrow(length=unit(0.3,'cm'))) +
      geom_edge_loop(aes(y=edge.y,width=sqrt((E(g)$weight)), span=50,direction=0,strength=0.90), color="grey85") +
      geom_node_point(aes(y=node.y),color=factor(V(g)$color),
                      size=12*sqrt(vsizes)/sqrt(max(vsizes))) +
      geom_node_text(aes(y=node.y.lab,label = common_name), size=5) +
      geom_image(data=crab_node, aes(x=x,y=y,image=crab_image), size=0.07) +
      xlab(ylabel) +
      xlim(c(-2.7,2.7)) +
      # ylim(c(-0.09,1.15)) +
      theme_void() +
      theme(legend.position="none", 
            axis.title.y=element_text(angle=90,size=20,hjust=0.5),
            panel.border = element_rect(colour = "black", fill=NA, size=1)) +
      coord_flip() +
      scale_y_reverse(expand=expansion(add=c(0.15,1)))
  }
}
plot_directed_early(graphs_list[[1]])
```
<br>



